scaled k5 m5

### 2019-07-09 Papers/PartialShapeMatching/MissingShapes.tex
Updated files.

\documentclass[11pt]{amsart}
\usepackage{amsmath,amssymb,graphicx,color,setspace,enumerate,natbib}
\newcommand{\subscript}[2]{$#1#2$}
\newcommand{\Eyx}{\mathbb{E}_{Y|X}}
\newcommand{\Ex}{\mathbb{E}_X}
\newcommand{\sm}{\sup_{x \in M_c}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\C}{C}
\newcommand{\X}{X}
\newcommand{\CX}{\C^m \circ \X}
\newcommand{\Rtwo}{\mathbb{R}^2}
\newcommand{\ltwo}{\mathbb{L}^2}
\newcommand{\uc}{\mathbb{S}}
\DeclareMathOperator*{\argmin}{argmin}
%\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\theoremstyle{definition}
\newtheorem{remark}{Remark}
%\newtheorem*{proof}{Proof}
\begin{document}
\title{Classification of shapes}
%\author{Gregory J. Matthews, Kartik Barath, Sebastian Kurtek, Juliet Brophy, George Thiruvathukal, Ofer Harel, Grady Flanary}
\author{Gregory J. Matthews et. al. }
\date{\today}

\maketitle
\section{Introduction}
Statistical analysis of shape plays an important role in many areas of science such as biology (\cite{OHiggins1989}, \cite{OHigginsDryden1993}, \cite{GoodallLange1989}), chemistry (\cite{DrydenEtAl2007,CzogielEtAl2011}), medicine (\cite{Bookstein1996, BrignellEtAl2010}), bioinformatics (\cite{GreenAndMardia2006}), genetics (\cite{HorganEtAl1992}), geology (\cite{Lohman1983}) and anthropology \cite{Brophy2014}(Matthews 2017).  Formal techniques for statistical shape analysis have been developed to extend classical statistical methods to shapes, such as computation of a mean shape and shape variability, extensions of Hotelling's $T^2$-test for inference about mean shapes, and principal components analysis of shapes (PCA), to name just a few \cite{DrydenAndMardiaBook}.  Statistical shape analysis techniques generally assume that the entire shape is observed.  However, in practice, there are instances where the shapes of interest are only partially observed.  For example, the shape of the occlusal surface of isolated fossil teeth from the Family Bovidae is useful to biological anthropologists for identifying the taxa of specimens, which in turn is used to reconstruct paleoenvirnoments.  In order to classify these isolated fossil tooth remains, scientists rely on, among other factors, the shape of the outline of the occlusal surface of the tooth to make accurate taxonomic classifications. Currently, only complete (or nearly complete) teeth are generally considered in classification analyses, however, there are numerous specimens of fractured fossil Bovid teeth that are not considered as strongly or ignored entirely in the analysis process because the full shape is not observed.  There are countless examples in the missing data literature where dropping observations with missing data (i.e. complete case analysis) has been demonstrated to produce biased results, therefore, it is scientifically beneficial to have methods for including partial observations in an analysis of interest, in this example the classification of fossilized teeth from the family Bovidae.

\subsection{Statistical Shape Analysis}
\subsection{Landmarks}
Much of the early work in shape analysis was based on defining shapes by a series of landmarks (\cite{Bookstein1986,MardiaAndDryden1989,Kendall1984,LeAndKendall1993}) where each point represents a meaningful location on a shape and there is a natural correspondence between points across shapes in a sample.  This allows for a natural correspondence of points across shapes and greatly simplifies analysis in many cases, such as computing a mean shape or variability of a collection of shapes.  The use of landmark data also allows for a natural methods for the alignment of shapes to one another using a method called Procrustes analysis \cite{Green1952,Goodall1991}, which aligns shapes subject to translations, rotations, and scaling.  Additionally, shapes can be defined by unlabeled points where there is some true labeling, but it is unobserved.  This setting is more complicated than labeled landmarks since the correspondence between points across shapes is unknown.  In many situations, however, there are no reasonable choices for landmarks on the shape.  Later,the concept of sliding landmarks \citet{Green1996, Bookstein1997} was proposed.  Sliding landmarks are the same as landmarks except that they are allowed to ``slide" along the contour of the curve when being aligned with a reference set of landmarks.

\subsection{Continuous Curves1}
Rather than describing a closed shape with a discrete set of points, a function can be used: 
$$
\beta:  \mathbb{S} \rightarrow \mathbb{R}^n 
$$
where $\mathbb{S}$ is the unit sphere and guarantees that the function has the same start and end point (i.e. is a closed curve).  The function $\beta$ is a closed curve but scalings, translations, rotations and reparameterizations of this function will usually result in different functions, which means that this framework does not offer invariance across these transformations.  Rather than using this function directly, the first derivative could be used, which results in a translation invariant representation.

This representation is the foundation of the square root velocity function (SRVF) (CITE) framework, which allows invariance across tranlsation, rotation, (optional) scaling, and reparameterization.  Frameworks for representing shapes that have invariance over reparameterization are said to be {\em elastic}.

To define the SRVF start by considering an absolutely continuous function $\beta:  \mathbb{S} \rightarrow \mathbb{R}^n$.  The SRVF of this function $q:  \mathbb{S} \rightarrow \mathbb{R}^n$ is defined as the following:

%Need to add a dot here.  
% $$
% q(t) = \left{ \begin{array}{ll}
%         \frac{\beta(t)}{\sqrt{||\beta(t)||}} & \mbox{If \beta(t) exist and is nonzero;}\\
%         0 & otherwise \end{array} \right
% $$

where $\dot{\beta(t)}$ is the first derivative of the function $\beta(t)$.

A full review of other methods for shape analysis can be found in CITE(FSDA)

%Statistical shape analysis is interested in studying a wide variety of questions about shapes from statistical view such as, for example, how different are two shapes, what is the mean / variability of a collection of shapes, and, of particular interest in this manuscipt, shape classification.  In order to work with shapes statistically, a framework needs to be defined for representing shapes, and there are many different idea for doing this.  For instance, shapes can be represented by landmarks \citep{Kendall1984,LeAndKendall1993,MardiaAndDryden1989}. A landmark is a point on a shape that corresponds to a point on another shape across and within populations.

Sliding landmarks by Bookstein.

Section 1.2.4 in FSU disserttion is a good description of this.

Shape analysis in general.

Partial Shape matching. 
%http://diginole.lib.fsu.edu/islandora/object/fsu%3A183299

\section{Missing Data}
Missing data problems arise in many statistical settings due to survey non-response, designed missingness in surveys (CITE), measurement error, and latent classes, to name a few examples.  There are numerous methods proposed in the literaure describing methods for handling missing data in a statistically principled manner.  For example, (Demspter Laird and Rubin) describes the EM algorithm for finding maximum likelihood estimates in the presence of missing data.  Another method, of particular interest in this setting, for handling mising data is multiple imputation (CITE Rubin and Little, Schafer).  The idea behind multiple imputation is to blah blah blah.......

EM algorithm (Depster Laird and Rubin)
Multiple Imputation (Rubin and Little)

For a more thorough review of the missing data literature, see cite(HortonAndKleinman2007) (Add more).

\section{Missing Data and Shapes}

\subsection{13.1 - Incomplete Data}
Papers mentioned: 
\begin{itemize}
\item Albers and Gower (2010) - Missing data in procrustes
\item Mitchelson (2013) - handling occluded landmarks. 
\item Gunz et. al. (2009) - missing data in studies of biological evolution of bone surfaces
\end{itemize}

\section{Shapes of closed curves1}
Our interest is in the completion of shapes of partially observed closed curves, and their classification. This first requires us to adopt a suitable representation for the shape of a fully observed curve. We adopt a parametric representation of a closed curve by representing as an absolutely continuous function\footnote{Upon identification of $\uc$ with $[0,1] \cong \mathbb{R}/2\pi\mathbb{Z}$, a curve $\C:[0,1] \to \mathbb{R}^2$ is absolutely continuous if and only if there exists an integrable function $g:[0,1]\to \Rtwo$ such that $\C(t)-\C(0)=\int_0^tg(u)du, \forall t \in [0,1]$.} $\C: \uc \to \Rtwo$, thus automatically ensuring that the curve is closed. The notion of a shape of such a curve requires invariances to transformations that represent nuisance information. Specifically, if $\Gamma:=\{\gamma:S \to S \text{ is an orientation-preserving diffeomorphism}\}$ and $SO(2)$ is the rotation group in $\mathbb{R}^3$, the shape of a parameterized curve $\C:D \to \Rtwo$ is defined to be the equivalence class 
\[
[\C]:=\Big\{\sigma O \C(\gamma(t))+a, \gamma \in \Gamma, O \in SO(2), a \in \Rtwo, \sigma >0\Big\}.
\]
Thus $[\C]$ is the set of all possible curves that can be obtained through a translation ($\C+a$), rotation ($O\C$), scale change $(\sigma \C)$, a reparametrization ($C(\gamma)$), or a combination of the transformations, of the curve $C$. In words, the shape of a curve $C$ is what is left once variations due scale, translation, rotation and reparametrisation has been accounted for.

In this instance, it is also of interest to preserve the size and shape of the curve.  This simply requires removing any scale changes from the definition of the equivalence class.  This size-and-shape of a parameterized $C: D \rightarrow \mathbb{R}^2$ is defined to be the equivalence class:

\[
[\C^{\star}]:=\Big\{\sigma O \C(\gamma(t))+a, \gamma \in \Gamma, O \in SO(2), a \in \Rtwo, \Big\}.
\]

Thus $[C^{\star}]$ is the set of all possible curves that can be obtained through a translation ($C+a$), rotation ($OC$), a reparameterization ($C(\gamma)$), or any combination of the transofrmations of curve $C$, notably with scaling not included as one of the possible transformations.

@@

\subsection{Notation}
A closed planar curve $C$ is always an absolutely continuous mapping $C:\uc \to \Rtwo$, and the set of planar closed curves will be denoted by $\mathcal{C}$. The set $\mathcal{C}$ is equipped with the norm $\|x\|_{\ltwo}:=[\int_\uc \|x(t)\|_2dt]^{1/2}$ where $\|\cdot\|_2$ is the Euclidean norm in $\Rtwo$. $SO(2)$ is the special orthogonal group of rotation matrices of $\Rtwo$, and $\Gamma$ is group of orientation-preserving diffeomorphisms of $\uc$. The set $\Gamma_I$ will denote the group $\{\gamma:[0,1]\to[0,1], \gamma'>0,\gamma(0)=0, \gamma(1)=1\}$.

\subsection{Square-root velocity transform}
For a detailed introduction to the transform, its properties and advantages we refer to the reader to Chapter 6 of the book by \cite{AK}. Here we briefly outline the important concepts required for our purposes.

For an absolutely continuous curve $\C:\uc \to \Rtwo$ consider the transformation 
$$\C \mapsto \frac{C'}{\|C'\|_{\ltwo}}=:Q_\C,$$ 
where $\C'$ is the (vector) derivative of $\C(t)$ with respect to $t$, and $\|\cdot\|$ is the usual Euclidean norm in $\Rtwo$.

@@

Two important ramifications of the described framework, motivating its use in our work for analyzing shapes of curves, are the following. Under the square-root velocity framework:
\begin{enumerate}[1.]
\item The actions of $SO(2)$ and $\Gamma$ on $\mathcal{Q}$ commute, i.e. they can be applied to a curve in any order. This ensures that their combined action is given by the product group $\Gamma \times SO(2)$.
\item The action of $\Gamma \times SO(2)$ on $\mathcal{Q}$ is by isometries: Given two curves $\C_1$ and $\C_2$, we have $\|OQ_{\C_1}(\gamma)\sqrt{\gamma '}-OQ_{\C_2}(\gamma)\sqrt{\gamma '}\|_{\ltwo}=\|Q_{C_1}-Q_{C_2}\|_{\ltwo}$, for every $(\gamma,O) \in \Gamma \times SO(2)$. This ensures that if two square-root velocity transformed curves are rotated and reparameterized the same way, their distance remains unchanged. 
\end{enumerate}
 Starting with a curve $\C$ we can now define its shape to be the equivalence class or its orbit of its corresponding square-root transform:
 $$[Q_\C]=\text{closure}\{OQ_\C(\gamma) \sqrt{\gamma'}: (\gamma,O) \in \Gamma \times SO(2)\},$$
 where the closure is with respect to the norm $\|\cdot\|_{\ltwo}$ on $Q$. The shape space consequently is defined as $\mathcal{Q}_s:=\{[Q_\C]: Q_\C \in \mathcal{Q}\}$. Property (2) ensures that the metric induced by the norm $\|\cdot\|_{\ltwo}$ on $\mathcal{Q}$ descends onto a metric $d$ on the shape space (quotient space) $\mathcal{Q}_s$ in a natural way. Given two curves $\C_1$ and $\C_2$, the shape distance between them is defined as
 \begin{align}
 \label{distance}
 d(\C_1,\C_2)&:=\inf_{(\gamma,O) \in \Gamma \times SO(2)} \|Q_{\C_1} - OQ_{\C_2}(\gamma)\sqrt{\gamma'}\|_{\ltwo}\\
 &=\inf_{(\gamma,O) \in \Gamma \times SO(2)}\Big[\int_{\uc}\left\|Q_{C_1}(t)-OQ_{C_2}(\gamma(t))\sqrt{\gamma'(t)}\right\|_2^2dt\Big]^{1/2} \nonumber\\
 &=\inf_{(\gamma,O) \in \Gamma \times SO(2)} \|OQ_{C_1}(\gamma)\sqrt{\gamma'} - Q_{C_2}\|_{\ltwo}\nonumber.
  \end{align}
  The symmetry with respect to the action either on $Q_{C_1}$ or $Q_{C_2}$ is an attractive feature and will be used profitably in the sequel.

\section{Preserving scale}
For an absolutely continuous curve $\C:\uc \to \Rtwo$ consider the transformation 
$$\C \mapsto C'=:Q^{\star}_\C = Q_\C\|C'\|_{\ltwo} $$ 
where $\C'$ is the (vector) derivative of $\C(t)$ with respect to $t$.

In this case, the unique (again, up to translations) inverse of this curve $Q^{\star}_C$ is $\int_0^t Q^{\star}_C(s)ds$. To ensure that the curves are closed we need to impose an additional constraint: $\int_\uc Q^{\star}_C(t)=0$. 
  Thus the image of the set of absolutely continuous closed curves , $ \{\C:\uc \to \Rtwo: \int_\uc \|C'(t)\|_2dt=l\}$, under the scale preserving transform map $\C \mapsto Q^{\star}_\C$ is the set

\[
\mathcal{Q}:=\Big\{Q^{\star}_\C:  \int_\uc Q^{\star}_C(t)dt=0\Big\},
\]

 Starting with a curve $\C$ we can now define its size and shape to be the equivalence class or its orbit of its corresponding velocity transform:
 $$[Q^{\star}_\C]=\text{closure}\{OQ^{\star}_\C(\gamma) \sqrt{\gamma'}: (\gamma,O) \in \Gamma \times SO(2)\},$$
 where the closure is with respect to the norm $\|\cdot\|_{\ltwo}$ on $Q^{\star}$.

 In the square-root velocity framework, dividing the velocity function by its norm gives us a function that is of norm one.  This implies that it is a point on the unit sphere.  In the finite-dimensional vecotr set-up this amounts to keeping just the direction of the vectors and discarding all other information and the distance between two curves can then be viewed as the angle between these two vectors.  However, when preserving scale this defintion of distance will need to be modified as both the angle AND magnitude of the vectors are both of interest.

 Therefore, a modified defintion of distance that preserves scaling between two curves will be defined:

 %Is this one symmetric?  I think so.  
 \begin{align}
 \label{distancestar}
 d^{\star}(\C_1,\C_2)&:=\inf_{(\gamma,O) \in \Gamma \times SO(2)} \|Q^{\star}_{\C_1} - OQ^{\star}_{\C_2}(\gamma)\sqrt{\gamma'}\|_{\ltwo}\\
 &=\inf_{(\gamma,O) \in \Gamma \times SO(2)}\Big[\int_{\uc}\left\|Q^{\star}_{C_1}(t)-OQ^{\star}_{C_2}(\gamma(t))\sqrt{\gamma'(t)}\right\|_2^2dt\Big]^{1/2} \nonumber\\
 &=\inf_{(\gamma,O) \in \Gamma \times SO(2)} \|OQ^{\star}_{C_1}(\gamma)\sqrt{\gamma'} - Q^{\star}_{C_2}\|_{\ltwo}\nonumber.
  \end{align}

 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Missingness Function}
Now consider a function, $R$, defined over the set $\mathbb{S}$, where $R(t) = 1$ if the shape function is observed at $t \in \mathbb{S}$ and $R(t) = 0$ otherwise.

Specifically,   
$$
R: \mathbb{S} \rightarrow \left\{0,1\right\}^2
$$

In $\Rtwo$, for a given point $t$ we have $C(t) = (x_t, y_t)$ and the defintion of the function $R$ allows these two be missing individually (e.g. $x_t$ could be observed and $y_t$ is missing or $y_t$ could be observed and $x_t$ is missing), in our setting both the $x_t$ and $y_t$ values are either both observed or both missing and so specifically in this case the function is of the following form: 
$$
R: \mathbb{S} \rightarrow \left\{\{0,0\},\{1,1\}\right\}
$$

Now for a specific curve $C$ in the equivalence class $[\C]$ (or $[\C^{\star}]$), we can define the function $R$ over $\mathbb{S}$.  However, for every reparameterization of $C$ in the equivalence class $[\C]$ (or $[\C^{\star}]$), there is a corresponding reparameterization of $R$.  Also, note that $R^{-1}(1)$ is the subset of domain $\mathbb{S}$ such that the coordinates of the shape are observed for every point in this set.  Further, here we only the case where $R^{-1}(1)$ be a connected subset of $\mathbb{S}$ is considered, though in general, this does not have to be the case. This prevents us from considering partially observed shapes with multiple ``gaps" (though this is an interesting idea for future work).

For a given shape, there are many possible reparameterizations of the missingness function, and an equivalence class can be defined for the missingness function across reparametertizations as follows:

%I think I only need reparameterizations for R?  Rotations too?  
$$
[\R] = \{ R(\gamma(t)) , \gamma \in \Gamma \} 
$$.

Note that for every curve $\C$ and corresponding $R$, in order to maintain the correct relationship, if $\C$ is reparameterized by $\gamma \in \Gamma$ then $R$ must be reparameterized using the same $\gamma$ that acted on $\C$.

Given $\R$, we can break $[\C^{\star}]$ into the observed and missing part of the size and shape as follows:

$$
[\C^{\star}]_{obs} = \{ O C(\gamma(t)) + a,  R(\gamma(t)), \gamma \in \Gamma, O \in SO(2), a\in \mathbb{R}^2 |  R(\gamma(t)) = 1  \}
$$

and
%Could also say where $t \in \gamma^{-1}(R^{-1}(1))$

$$
[\C^{\star}]_{mis} = \{ O C(\gamma(t)) + a,  R(\gamma(t)), \gamma \in \Gamma, O \in SO(2), a\in \mathbb{R}^2 |  R(\gamma(t)) = 0  \}
$$.

Likewise for $[\C]$ we can do the same thing, but also consider scaling: 
$$
[\C]_{obs} = \{\sigma O C(\gamma(t)) + a,  R(\gamma(t)), \gamma \in \Gamma, O \in SO(2), a\in \mathbb{R}^2, \sigma \> 0 |  R(\gamma(t)) = 1  \}
$$

and
%Could also say where $t \in \gamma^{-1}(R^{-1}(1))$

$$
[\C]_{mis} = \{\sigma O C(\gamma(t)) + a,  R(\gamma(t)), \gamma \in \Gamma, O \in SO(2), a\in \mathbb{R}^2, \sigma \> 0 |  R(\gamma(t)) = 0  \}
$$.

\subsection{Completing the curve}
%lets use partial and full to denote which curve is which.  Then the observed part of \C^p_j is also \C^p_{j,obs} but there is also a \C^p_{j,mis}, which is the missing part.  Then the concatenation of \C^p_{j,obs} and \C^p_{j,mis} is \C^p_j.

Suppose we are given closed planar curves $\C^p_j:\uc \to \mathbb{R}^2,j=1\ldots,n_p$ each of which has been observed only on a region $\{t : R(\gamma(t) = 1\} \subset \uc$, where $\uc$ is the unit circle in $\Rtwo$. We assume that the curves are absolutely continuous. Additionally, a training sample $\{(y_i,\C^f_i),1,\ldots,n_f\}$ consisting of class labels $y_i \in \{0,1, \cdots G\}$ and $C^f_i:\uc\to \mathbb{R}^2$, fully observed on a common domain $\uc$, is provided. The set of fully observed curves are elements of the set $\mathcal{C}^f$; denote by $\mathcal{C}^p$ the set of all partially observed curves.

%I don't want to get too hung up on classification.  Classification is just one thing you could do after completing the curves.  
The problem at hand is to model the shape (and size) of and complete each partially observed curve $\C^p_j$ and assign it to one of $G$ groups. We consider two approaches: one based on a variational formulation, and the other using kernel-based classifier. Before describing our approaches, some comments on the set $\Gamma$ are in order.

@@

Consider now the set 
$$
\Gamma_\mathbb{R}:=\{\beta:\mathbb{R}\to \mathbb{R}: \beta(t+1)=\beta(t)+1, \text{ continuous and increasing}\}
$$.

Each member $\beta$ of $W_\mathbb{R}$ induces a warp map $\tilde{\beta}:\uc \to \uc$ with $\tilde{\beta}(e^{2\pi it})=e^{2\pi i\beta(t)}$, where $\beta$ is referred to as the lift of $\tilde{\beta}$. This $\beta$ satisfies $\beta(t+1)=\beta(t)+1$ for all $t \in [0,1]$, and consequently we have, for $t \in [0,1]$, $\beta(t)=\gamma(t)+c $, where $\gamma$ is a warp map of $[0,1]$ and $c \in (0,1]$ (through the identification of $[0,1]$ with $\mathbb{R}/2\pi \mathbb{Z}$). This procedure can be viewed as one that produces a warp map of $\uc$ by `unwrapping' $\uc$ at a chosen point $s$ and generating a warp map of $[0,1]$.
If $\Gamma_I:=\{\gamma:[0,1] \to [0,1]: \gamma'>0,\gamma(0)=0, \gamma(1)=1\}$ is the group of diffeomorphisms of $[0,1]$, the map $\Gamma \mapsto \uc \times \Gamma_I$ is a bijection \footnote{Technically, this is not a bijection since for a $\gamma \in \Gamma$, the corresponding $\beta \in \Gamma_I$ has a jump discontinuity at the point $t_c \in [0,1]$ where $\beta(t_c)+c=1$. This can be circumvented by assuming that the members of $\Gamma$ and $\Gamma_I$ are absolutely continuous (as opposed to diffeomorphims); then the map between to the sets is bijective a.e.}. We will hence employ the product group $\uc \times \Gamma_I$ in place of $\Gamma$. This ensures that the domain of each curve $C_j^p$ and $C^f_i$ can be identified with $[0,1]$ upon unwrapping the circle. 
%This results in $\C^f_i$ becoming periodic with period 1: $\C^f_i(0)=\C^f_i(1)$ for every $i=1,\ldots,n_p$.

\subsection{Curve Completion1}
Out goal is to complete $[\C]_{obs}$ or ($[\C^{\star}]_{obs}$) by randomly drawing from the distribution on the missing part of the shape conditional on the observed part.  Specifically, for a fixed translation, rotation, scaling, and parameterization, we seek to approximate the distribution of $\C_{mis}|\C_{obs}$ and use random draws from this distribution to complete to curve.  Here this distribution will be approximated non-parametrically and hot-deck type multiple imputation \cite{AndridgeLittle2010} will be used similar in spirit to the idea of predictive mean matching (PMM) (\cite{PMM}).  The basic idea of PMM is to to take partially observed records (termed the donees) and find fully observed records that are close to a particular donee (termed donors).  Random draws from the set of donors are then used to complete the partially observed donees.  This is repeated $M$ times to create $M$ completed data sets.

The same type of procedure will be performed here, but instead of traditional statistical records, here shapes are considered.  The spirit of the procedure proposed here is similar to PMM in that first the partially observed shape (the donee) is measured against fully observed shapes to find potential donors.  Once the set of potential donors is defined, one of these shapes is randomly drawn and used to complete the partially observed donee.  This process is crepeated $M$ times to create $M$ completed shapes.

>>>>

\documentclass[11pt]{amsart}
\usepackage{amsmath,amssymb,graphicx,color,setspace,enumerate,natbib}
\newcommand{\subscript}[2]{$#1#2$}
\newcommand{\Eyx}{\mathbb{E}_{Y|X}}
\newcommand{\Ex}{\mathbb{E}_X}
\newcommand{\sm}{\sup_{x \in M_c}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\C}{C}
\newcommand{\X}{X}
\newcommand{\CX}{\C^m \circ \X}
\newcommand{\Rtwo}{\mathbb{R}^2}
\newcommand{\ltwo}{\mathbb{L}^2}
\newcommand{\uc}{\mathbb{S}}
\DeclareMathOperator*{\argmin}{argmin}
%\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\theoremstyle{definition}
\newtheorem{remark}{Remark}
%\newtheorem*{proof}{Proof}
\begin{document}
\section{Shapes of closed curves}
Our interest is in the completion of shapes of partially observed closed curves, and their classification. This first requires us to adopt a suitable representation for the shape of a fully observed curve. We adopt a parametric representation of a closed curve by representing as an absolutely continuous function\footnote{Upon identification of $\uc$ with $[0,1] \cong \mathbb{R}/2\pi\mathbb{Z}$, a curve $\C:[0,1] \to \mathbb{R}^2$ is absolutely continuous if and only if there exists an integrable function $g:[0,1]\to \Rtwo$ such that $\C(t)-\C(0)=\int_0^tg(u)du, \forall t \in [0,1]$.} $\C: \uc \to \Rtwo$, thus automatically ensuring that the curve is closed. The notion of a shape of such a curve requires invariances to transformations that represent nuisance information. Specifically, if $\Gamma:=\{\gamma:S \to S \text{ is an orientation-preserving diffeomorphism}\}$ and $SO(2)$ is the rotation group in $\mathbb{R}^3$, the shape of a parameterized curve $\C:D \to \Rtwo$ is defined to be the equivalence class 
\[
[\C]:=\Big\{\sigma O \C(\gamma(t))+a, \gamma \in \Gamma, O \in SO(2), a \in \Rtwo, \sigma >0\Big\}.
\]
Thus $[\C]$ is the set of all possible curves that can be obtained through a translation ($\C+a$), rotation ($O\C$), scale change $(\sigma \C)$, a reparametrization ($C(\gamma)$), or a combination of the transformations, of the curve $C$. In words, the shape of a curve $C$ is what is left once variations due scale, translation, rotation and reparametrisation has been accounted for.

@@

\subsection{Notation}
A closed planar curve $C$ is always an absolutely continuous mapping $C:\uc \to \Rtwo$, and the set of planar closed curves will be denoted by $\mathcal{C}$. The set $\mathcal{C}$ is equipped with the norm $\|x\|_{\ltwo}:=[\int_\uc \|x(t)\|_2dt]^{1/2}$ where $\|\cdot\|_2$ is the Euclidean norm in $\Rtwo$. $SO(2)$ is the special orthogonal group of rotation matrices of $\Rtwo$, and $\Gamma$ is group of orientation-preserving diffeomorphisms of $\uc$. The set $\Gamma_I$ will denote the group $\{\gamma:[0,1]\to[0,1], \gamma'>0,\gamma(0)=0, \gamma(1)=1\}$. 
\subsection{Square-root velocity transform}
For a detailed introduction to the transform, its properties and advantages we refer to the reader to Chapter 6 of the book by \cite{AK}. Here we briefly outline the important concepts required for our purposes. 
For an absolutely continuous curve $\C:\uc \to \Rtwo$ consider the transformation 
$$\C \mapsto \frac{C'}{\|C'\|_{\ltwo}}=:Q_\C,$$ 
where $\C'$ is the (vector) derivative of $\C(t)$ with respect to $t$, and $\|\cdot\|$ is the usual Euclidean norm in $\Rtwo$.

@@

Two important ramifications of the described framework, motivating its use in our work for analyzing shapes of curves, are the following. Under the square-root velocity framework:
\begin{enumerate}[1.]
\item The actions of $SO(2)$ and $\Gamma$ on $\mathcal{Q}$ commute, i.e. they can be applied to a curve in any order. This ensures that their combined action is given by the product group $\Gamma \times SO(2)$.
\item The action of $\Gamma \times SO(2)$ on $\mathcal{Q}$ is by isometries: Given two curves $\C_1$ and $\C_2$, we have $\|OQ_{\C_1}(\gamma)\sqrt{\gamma '}-OQ_{\C_2}(\gamma)\sqrt{\gamma '}\|_{\ltwo}=\|Q_{C_1}-Q_{C_2}\|_{\ltwo}$, for every $(\gamma,O) \in \Gamma \times SO(2)$. This ensures that if two square-root velocity transformed curves are rotated and reparameterized the same way, their distance remains unchanged. 
\end{enumerate}
 Starting with a curve $\C$ we can now define its shape to be the equivalence class or its orbit of its corresponding square-root transform:
 $$[Q_\C]=\text{closure}\{OQ_\C(\gamma) \sqrt{\gamma'}: (\gamma,O) \in \Gamma \times SO(2)\},$$
 where the closure is with respect to the norm $\|\cdot\|_{\ltwo}$ on $Q$. The shape space consequently is defined as $\mathcal{Q}_s:=\{[Q_\C]: Q_\C \in \mathcal{Q}\}$. Property (2) ensures that the metric induced by the norm $\|\cdot\|_{\ltwo}$ on $\mathcal{Q}$ descends onto a metric $d$ on the shape space (quotient space) $\mathcal{Q}_s$ in a natural way. Given two curves $\C_1$ and $\C_2$, the shape distance between them is defined as
 \begin{align}
 \label{distance}
 d(\C_1,\C_2)&:=\inf_{(\gamma,O) \in \Gamma \times SO(2)} \|Q_{\C_1} - OQ_{\C_2}(\gamma)\sqrt{\gamma'}\|_{\ltwo}\\
 &=\inf_{(\gamma,O) \in \Gamma \times SO(2)}\Big[\int_{\uc}\left\|Q_{C_1}(t)-OQ_{C_2}(\gamma(t))\sqrt{\gamma'(t)}\right\|_2^2dt\Big]^{1/2} \nonumber\\
 &=\inf_{(\gamma,O) \in \Gamma \times SO(2)} \|OQ_{C_1}(\gamma)\sqrt{\gamma'} - Q_{C_2}\|_{\ltwo}\nonumber.
  \end{align}
  The symmetry with respect to the action either on $Q_{C_1}$ or $Q_{C_2}$ is an attractive feature and will be used profitably in the sequel. 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Curve completion and classification}
Suppose we are given closed planar curves $\C^m_j:\uc \to \mathbb{R}^2,j=1\ldots,n$ each of which has been observed only on a region $\R_j \subset \uc$, where $\uc$ is the unit circle in $\Rtwo$. We assume that the curves are absolutely continuous. Additionally, a training sample $\{(y_i,\C^o_i),1,\ldots,N\}$ consisting of class labels $y_i \in \{0,1\}$ and $C^o_i:\uc\to \mathbb{R}^2$, fully observed on a common domain $\uc$, is provided. The set of fully observed curves are elements of the set $\mathcal{C}$; denote by $\mathcal{C}^m$ the set of all partially observed curves.

The problem at hand is to model the shape of and complete each partially observed curve $\C^m_j$ and assign it to one of $G$ groups. We consider two approaches: one based on a variational formulation, and the other using kernel-based classifier. Before describing our approaches, some comments on the set $\Gamma$ are in order.

@@

Consider now the set $$\Gamma_\mathbb{R}:=\{\beta:\mathbb{R}\to \mathbb{R}: \beta(t+1)=\beta(t)+1, \text{ continuous and increasing}\}.$$ Each member $\beta$ of $W_\mathbb{R}$ induces a warp map $\tilde{\beta}:\uc \to \uc$ with $\tilde{\beta}(e^{2\pi it})=e^{2\pi i\beta(t)}$, where $\beta$ is referred to as the lift of $\tilde{\beta}$. This $\beta$ satisfies $\beta(t+1)=\beta(t)+1$ for all $t \in [0,1]$, and consequently we have, for $t \in [0,1]$, $\beta(t)=\gamma(t)+c $, where $\gamma$ is a warp map of $[0,1]$ and $c \in (0,1]$ (through the identification of $[0,1]$ with $\mathbb{R}/2\pi \mathbb{Z}$). This procedure can be viewed as one that produces a warp map of $\uc$ by `unwrapping' $\uc$ at a chosen point $s$ and generating a warp map of $[0,1]$.
If $\Gamma_I:=\{\gamma:[0,1] \to [0,1]: \gamma'>0,\gamma(0)=0, \gamma(1)=1\}$ is the group of diffeomorphisms of $[0,1]$, the map $\Gamma \mapsto \uc \times \Gamma_I$ is a bijection\footnote{Technically, this is not a bijection since for a $\gamma \in \Gamma$, the corresponding $\beta \in \Gamma_I$ has a jump discontinuity at the point $t_c \in [0,1]$ where $\beta(t_c)+c=1$. This can be circumvented by assuming that the members of $\Gamma$ and $\Gamma_I$ are absolutely continuous (as opposed to diffeomorphims); then the map between to the sets is bijective a.e.}. We will hence employ the product group $\uc \times \Gamma_I$ in place of $\Gamma$. This ensures that the domain of each curve $C_j^m$ and $C^o_i$ can be identified with $[0,1]$ upon unwrapping the circle. 
%This results in $\C^o_i$ becoming periodic with period 1: $\C^o_i(0)=\C^o_i(1)$ for every $i=1,\ldots,n$.

\subsection{Curve completion}
The observed region $\mathcal{R}_j$ associated with a partially observed curve $\C^m_j$ is the subinterval $[0,t_j]$ with $t_j<1$ for all $j=1,\ldots,N$. Suppose that $\C^m_j(0)=\mathbf{a}_j:=(a_{1j},a_{2j})^T$ and $\C^m_j(t_j)=\mathbf{b}_j:=(b_{1j},b_{2j})^T$. Then the set of curves comprising the missing segment of curve $\C^m_j$ is 
$$\mathcal{X}_j:=\{\X: [t_j,1] \to \Rtwo: \X(t_j)=\mathbf{b}, \X(1)=\mathbf{a}\} .$$
For a partially observed $\C_j^m:[0,t_j]$ and an $X \in \mathcal{X}_j$ with $j=1,\ldots,n$, define its completion to be the concatenated closed curve
\begin{equation*}
C_j \circ X(t):= \C_j^m(t) \mathbb{I}_{t \in [0,t_j]}+\X(t) \mathbb{I}_{(t_j,1]}.
\end{equation*}
